---
title: "Paser data analysis with big library"

output:
  html_document:
    anchor_sections: false
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
    fig_width: 11
    fig_height: 7
    code_folding: "hide"
    theme: united
toc-title: "Table of Contents"
---

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  #library(optparse)
  library(dplyr)
  library(tidyverse)
  library(tibble)
  library(ggplot2)
  library(RColorBrewer)
})
```

```{r set-parameters, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
paser_results_big <- "C:\\Users\\BhatA\\Box\\ASMS\\paser_big_lib.tsv"

paser_big_corr <- "C:\\Users\\BhatA\\Box\\ASMS\\sample_matrix_big_20on20_withCorr.tsv"
```

```{r read-input - small lib, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
sample_matrix_big <- read.delim(paser_results_big)

sample_matrix_big_copy <- sample_matrix_big

#sample_matrix_small <- filter(sample_matrix_small, Identified.By != "MBR")

sample_matrix_big$Protein <- paste(sample_matrix_big$Protein.Ids,sample_matrix_big$Precursor.Id,sep= "-")

sample_matrix_big <- sample_matrix_big[(colnames(sample_matrix_big) %in% c("Protein", "Precursor.Quantity","File.Name", "RT", "Exp.1.K0"))] 

sample_matrix_big$Category = str_split_fixed(sample_matrix_big$File.Name, '_', 6)[,5]

sample_matrix_big <- filter(sample_matrix_big, Category %in% c("A", "B", "C", "D"))

write.table(sample_matrix_big, file = "sample_matrix_big_noE.tsv", sep="\t", row.names=F, quote=F) # export to file


#sample_matrix_5on5 <- sample_matrix_small %>% group_by(Category, Protein) %>%
#  summarise(Precursor.Quantity = n())

#sample_matrix_5on5 <- filter(sample_matrix_5on5, Precursor.Quantity == 5)

#sample_matrix_small_filt_5on5 <- filter(sample_matrix_small,Protein %in% sample_matrix_5on5$Protein)

sample_matrix_big_cv <- sample_matrix_big %>% 
    group_by(Category,Protein) %>%
    summarise(StdDev = sd(Precursor.Quantity), Avg = mean(Precursor.Quantity), CV = (StdDev/Avg)*100)

# CV filter
sample_matrix_big_cv_filt <- filter(sample_matrix_big_cv, CV > 20)

# sample_matrix_small_20on20 is list with prec < CV 20
sample_matrix_big_20on20 <- filter(sample_matrix_big,!Protein %in% sample_matrix_big_cv_filt$Protein)

sample_matrix_big_20on20 <- sample_matrix_big_20on20 %>%
  group_by(Protein) %>%   
  summarise(Precursor.Quantity = n())

sample_matrix_big_20on20 <- filter(sample_matrix_big_20on20, Precursor.Quantity == 20)

sample_matrix_big_20on20_final <- filter(sample_matrix_big,Protein %in% sample_matrix_big_20on20$Protein)

write.table(sample_matrix_big_20on20_final, file = "sample_matrix_big_20on20_filtCV.tsv", sep="\t", row.names=F, quote=F) # export to file

```

## Precursor counts {.tabset}
```{r count-precursors, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# All data
summarize_precursor <- sample_matrix_big %>% group_by(Category) %>% summarise(Count = n_distinct(Protein))

# Filtered
summarize_precursor_filt <- sample_matrix_big_20on20_final %>% group_by(Category) %>% summarise(Count = n_distinct(Protein))

col_bar <- brewer.pal(4, "Blues") 

```

### Before filtering
```{r bar plot before filt, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

ggplot(data=summarize_precursor, aes(x=Count, y=Category, fill=Category)) +
  geom_bar(stat="identity") + 
  scale_fill_brewer(palette="Blues") + 
  labs(x = "Precursor Count",
      y = "Category") + 
  ggtitle("Precursor Count per Category") + 
  theme(plot.title = element_text(hjust = 0.5))

```

### After filtering
```{r bar plot after filt, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

ggplot(data=summarize_precursor_filt, aes(x=Count, y=Category, fill=Category)) +
  geom_bar(stat="identity") + 
  scale_fill_brewer(palette="Blues") + 
  labs(x = "Precursor Count",
      y = "Category") +
  ggtitle("Precursor Count per Category") + 
  theme(plot.title = element_text(hjust = 0.5))

```

## Precursor abundance (observed vs expected) {.tabset}
```{r precursor abundance, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# All data
median_grp_precursor = sample_matrix_big %>% group_by(Category, Protein) %>% summarise(Precursor.Quantity = median(Precursor.Quantity))

median_grp = median_grp_precursor %>% group_by(Category) %>%  summarise(Precursor.Quantity = median(Precursor.Quantity))

mean_grp = median_grp_precursor %>% group_by(Category) %>%  summarise(Precursor.Quantity = mean(Precursor.Quantity))

median_grp$median_percentage = (median_grp$Precursor.Quantity/median_grp$Precursor.Quantity[[1]]) * 100
median_grp$mean_percentage = (mean_grp$Precursor.Quantity/mean_grp$Precursor.Quantity[[1]]) * 100
#median_grp = head(median_grp, -1)
median_grp = median_grp[order(median_grp$Category, decreasing = TRUE),]
median_grp$expected = c(5, 10, 50, 100)
median_grp$scale = c(5, 10, 50, 100)

# Filtered
median_grp_precursor_filt = sample_matrix_big_20on20_final %>% group_by(Category, Protein) %>% summarise(Precursor.Quantity = median(Precursor.Quantity))

median_grp_filt = median_grp_precursor_filt %>% group_by(Category) %>%  summarise(Precursor.Quantity = median(Precursor.Quantity))

mean_grp_filt = median_grp_precursor_filt %>% group_by(Category) %>%  summarise(Precursor.Quantity = mean(Precursor.Quantity))

median_grp_filt$median_percentage = (median_grp_filt$Precursor.Quantity/median_grp_filt$Precursor.Quantity[[1]]) * 100
median_grp_filt$mean_percentage = (mean_grp_filt$Precursor.Quantity/mean_grp_filt$Precursor.Quantity[[1]]) * 100
median_grp_filt = median_grp_filt[order(median_grp_filt$Category, decreasing = TRUE),]
median_grp_filt$expected = c(5, 10, 50, 100)
median_grp_filt$scale = c(5, 10, 50, 100)
```

### Before filtering
```{r line graph before, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# All data
colors <- c("Expected" = "black", "Median %" = "red", "Mean %" = "steelblue")

ggplot(median_grp, aes(scale)) +
  scale_color_manual(values = colors) +
  geom_line(aes(y = expected, color = "Expected")) +
  geom_point(aes(y = expected, color = "Expected")) +
  geom_line(aes(y = median_percentage, color = "Median %")) +
  geom_point(aes(y = median_percentage, color = "Median %")) +
  geom_line(aes(y = mean_percentage, color = "Mean %")) +
  geom_point(aes(y = mean_percentage, color = "Mean %")) +
  scale_x_discrete(limits=c(5,10,50,100)) +
  expand_limits(x=0, y=0) +
  #scale_x_continuous(limits=c(0,100)) +
  scale_y_continuous(limits=c(0,100), breaks = scales::breaks_width(10)) +
  labs(x = "Concentration",
        y = "Precursor abundance (%)",
        color = "Legend") + 
  ggtitle("Precursor abundance across concentrations") +
    theme(plot.title = element_text(hjust = 0.5))
```

### After filtering
```{r line graph after, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# Filtered
colors <- c("Expected" = "black", "Median %" = "red", "Mean %" = "steelblue")

ggplot(median_grp_filt, aes(scale)) +
  scale_color_manual(values = colors) +
  geom_line(aes(y = expected, color = "Expected")) +
  geom_point(aes(y = expected, color = "Expected")) +
  geom_line(aes(y = median_percentage, color = "Median %")) +
  geom_point(aes(y = median_percentage, color = "Median %")) +
  geom_line(aes(y = mean_percentage, color = "Mean %")) +
  geom_point(aes(y = mean_percentage, color = "Mean %")) +
  scale_x_discrete(limits=c(5,10,50,100)) +
  expand_limits(x=0, y=0) +
  #scale_x_continuous(limits=c(0,100)) +
  scale_y_continuous(limits=c(0,100), breaks = scales::breaks_width(10)) +
  labs(x = "Concentration",
        y = "Precursor abundance (%)",
        color = "Legend") + 
  ggtitle("Precursor abundance across concentrations") +
    theme(plot.title = element_text(hjust = 0.5))

```

## Coefficient of Variation {.tabset}
```{r CV, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# All data
cv_precursor <- sample_matrix_big %>% 
    group_by(Category,Protein) %>%
    summarise(StdDev = sd(Precursor.Quantity), Avg = mean(Precursor.Quantity), CV = StdDev/Avg*100)

# Filtered
cv_precursor_filt <- sample_matrix_big_20on20_final %>% 
    group_by(Category,Protein) %>%
    summarise(StdDev = sd(Precursor.Quantity), Avg = mean(Precursor.Quantity), CV = (StdDev/Avg)*100)
```

### Before filtering
```{r before filt CV, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

p <- ggplot(cv_precursor, aes(x=Category, y=CV, fill=Category)) +
    geom_violin() + 
    scale_fill_brewer(palette="Blues") + 
    theme_classic() +
    stat_summary(fun.data="mean_sdl", mult=1, 
                 geom="pointrange", width=0.2 ) + 
    ggtitle("CV % per Category") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    scale_y_continuous(limits=c(0,200), breaks = scales::breaks_width(20)) 
p

```

### After filtering
```{r after filt CV, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

p <- ggplot(cv_precursor_filt, aes(x=Category, y=CV, fill=Category)) +
    geom_violin() + 
    scale_fill_brewer(palette="Blues") + 
    theme_classic() +
    stat_summary(fun.data="mean_sdl", mult=1, 
                 geom="pointrange", width=0.2 ) + 
    ggtitle("CV % per Category") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    scale_y_continuous(limits=c(0,200), breaks = scales::breaks_width(20)) 
p

```

## Difference in distribution of Retention Time {.tabset}
```{r RT, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# All data
rt_precursor <- sample_matrix_big %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(RT))

# Filtered
rt_precursor_filt <- sample_matrix_big_20on20_final %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(RT))
```

### Before filtering
```{r before filtering RT, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
for (i in c("B", "C", "D")) {
  sub_A <- filter(rt_precursor, Category == "A")
  sub_i <- filter(rt_precursor, Category == i)
  
  merged_A_i <- merge(x = sub_A, y = sub_i, by = "Protein")
  merged_A_i <- merged_A_i %>% mutate(diff_iA = Median.y-Median.x)
  
  p <- ggplot(merged_A_i, aes(diff_iA)) +
  geom_histogram(binwidth = 0.01, color="steelblue", fill="steelblue") +
  ggtitle(paste0("RT difference between ", i, " and A")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") + 
  ylab("Frequency") + 
  xlim(-1.5, 1.5)
  
  print(p)
}
```

### After filtering
```{r after filetring RT, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
for (i in c("B", "C", "D")) {
  sub_A <- filter(rt_precursor_filt, Category == "A")
  sub_i <- filter(rt_precursor_filt, Category == i)
  
  merged_A_i <- merge(x = sub_A, y = sub_i, by = "Protein")
  merged_A_i <- merged_A_i %>% mutate(diff_iA = Median.y-Median.x)
  
  p <- ggplot(merged_A_i, aes(diff_iA)) +
  geom_histogram(binwidth = 0.01, color="steelblue", fill="steelblue") +
  ggtitle(paste0("RT difference between ", i, " and A")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") + 
  ylab("Frequency") + 
  xlim(-1.5, 1.5)
  
  print(p)
}

```

## Difference in distribution of Ion Mobility {.tabset}
```{r IM, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
#All data
im_precursor <- sample_matrix_big %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(Exp.1.K0))

# Filtered
im_precursor_filt <- sample_matrix_big_20on20_final %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(Exp.1.K0))
```

### Before filtering
```{r before filtering IM, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
for (i in c("B", "C", "D")) {
  sub_A <- filter(im_precursor, Category == "A")
  sub_i <- filter(im_precursor, Category == i)
  
  merged_A_i <- merge(x = sub_A, y = sub_i, by = "Protein")
  merged_A_i <- merged_A_i %>% mutate(diff_iA = Median.y-Median.x)
  
  p <- ggplot(merged_A_i, aes(diff_iA)) +
  geom_histogram(binwidth = 0.0001, color="steelblue", fill="steelblue") +
  ggtitle(paste0("IM difference between ", i, " and A")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") + 
  ylab("Frequency") + 
  xlim(-0.5, 0.5)
  
  print(p)
}
```

### After filtering
```{r after filtering IM, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
for (i in c("B", "C", "D")) {
  sub_A <- filter(im_precursor_filt, Category == "A")
  sub_i <- filter(im_precursor_filt, Category == i)
  
  merged_A_i <- merge(x = sub_A, y = sub_i, by = "Protein")
  merged_A_i <- merged_A_i %>% mutate(diff_iA = Median.y-Median.x)
  
  p <- ggplot(merged_A_i, aes(diff_iA)) +
  geom_histogram(binwidth = 0.0001, color="steelblue", fill="steelblue") +
  ggtitle(paste0("IM difference between ", i, " and A")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") + 
  ylab("Frequency") + 
  xlim(-0.5, 0.5)
  
  print(p)
}

```

## Distribution of Precursors based on correlation {.tabset}
```{r prec with good correlation, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

paser_big_corr_file <- read.delim(paser_big_corr)

# Subset to get good, bad and ugly precursors based on correlation
paser_big_corr_good <- filter(paser_big_corr_file, Corr >= 0.9)
paser_big_corr_bad <- filter(paser_big_corr_file, between(Corr, 0.5, 0.7))
paser_big_corr_ugly <- filter(paser_big_corr_file, between(Corr, 0.0, 0.4))

```

### Good
```{r Good, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

ggplot(paser_big_corr_good, aes(x=Category, y=log2(Precursor.Quantity), fill=Category)) +
  geom_boxplot() +
  scale_fill_brewer(palette="Blues") +
  theme(
    legend.position="none",
    plot.title = element_text(hjust = 0.5)
  ) +
  ggtitle("Distribution of Precursors across Concentrations - Good") +
  labs(x = "Concentrations",
      y = "Precursor abundance (log2)")

```

### Bad
```{r prec with bad correlation, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

ggplot(paser_big_corr_bad, aes(x=Category, y=log2(Precursor.Quantity), fill=Category)) +
  geom_boxplot() +
  scale_fill_brewer(palette="Blues") +
  theme(
    legend.position="none",
    plot.title = element_text(hjust = 0.5)
  ) +
  ggtitle("Distribution of Precursors across Concentrations - Bad") +
  labs(x = "Concentrations",
      y = "Precursor abundance (log2)")

```

### Ugly
```{r prec with ugly correlation, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

ggplot(paser_big_corr_ugly, aes(x=Category, y=log2(Precursor.Quantity), fill=Category)) +
  geom_boxplot() +
  scale_fill_brewer(palette="Blues") +
  theme(
    legend.position="none",
    plot.title = element_text(hjust = 0.5)
  ) +
  ggtitle("Distribution of Precursors across Concentrations - Ugly") +
  labs(x = "Concentrations",
      y = "Precursor abundance (log2)")

```
