---
title: "Paser data analysis with small library"

output:
  html_document:
    anchor_sections: false
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
    fig_width: 11
    fig_height: 7
    code_folding: "hide"
    theme: united
toc-title: "Table of Contents"
---

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  #library(optparse)
  library(dplyr)
  library(tidyverse)
  library(tibble)
  library(ggplot2)
  library(RColorBrewer)
  library(Hmisc)
})
```



```{r set-parameters, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# Results file
paser_results_small <- "Enter the path of your downloaded paser results small"

# Results file subset with correlation
paser_small_corr <- "Enter the path of your downloaded correlation file"
```


```{r set-parameters, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# RUN THIS SNIPPET OF CODE ONLY IF YOU'VE DOWNLOADED THE RDS FILES

#small library results file
sample_matrix_small <- readRDS('../data/sample_matrix_small.rds')

#results file subset with correlation
paser_small_corr_file <- readRDS('../data/paser_small_corr.rds')


```



```{r read-input - small lib, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# Read the paser results file
sample_matrix_small <- read.delim(paser_results_small)

#read in the correlation matrix file
paser_small_corr_file <- read.delim(paser_small_corr)

# Make a copy of the same data
#sample_matrix_small_copy <- sample_matrix_small
```


```{r read-input - small lib, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

# Remove precursors identified by MBR
sample_matrix_small <- filter(sample_matrix_small, Identified.By != "MBR")

# Concatenate protein and precursor
sample_matrix_small$Protein <- paste(sample_matrix_small$Protein.Ids,sample_matrix_small$Precursor.Id,sep= "-")

# Keep only columns we are interested in 
sample_matrix_small <- sample_matrix_small[(colnames(sample_matrix_small) %in% c("Protein", "Precursor.Quantity","File.Name", "RT", "Exp.1.K0"))] 

# Create Category column depending what kind of sample it is, A,B,C,Dor E!
sample_matrix_small$Category = str_split_fixed(sample_matrix_small$File.Name, '_', 6)[,5]

# Remove precursors from Category "E" since it contains no Hela
sample_matrix_small <- filter(sample_matrix_small, Category %in% c("A", "B", "C", "D"))

# Create a table with subset data, No MBR, No E. This is the file used to make plots under "All data".
#write.table(sample_matrix_small, file = "sample_matrix_small_noE.tsv", sep="\t", row.names=F, quote=F) # export to file

#sample_matrix_5on5 <- sample_matrix_small %>% group_by(Category, Protein) %>%
#  summarise(Precursor.Quantity = n())

#sample_matrix_5on5 <- filter(sample_matrix_5on5, Precursor.Quantity == 5)

#sample_matrix_small_filt_5on5 <- filter(sample_matrix_small,Protein %in% sample_matrix_5on5$Protein)

# Keep only precursors that are present in all 20 samples
sample_matrix_small_20on20 <- sample_matrix_small %>%
  group_by(Protein) %>%   
  summarise(Precursor.Quantity = n())

sample_matrix_small_20on20 <- filter(sample_matrix_small_20on20, Precursor.Quantity == 20)

sample_matrix_small_20on20_final <- filter(sample_matrix_small,Protein %in% sample_matrix_small_20on20$Protein)

# Create a table with filtered data, < 20% CV, No E. This is the file used to make plots with CV < 20 % and 20on20
#write.table(sample_matrix_small_20on20_final, file = "sample_matrix_small_filt20on20.tsv", sep="\t", row.names=F, quote=F) # export to file

# Calculate CV per Category and Precursor
sample_matrix_small_cv <- sample_matrix_small %>% 
    group_by(Category,Protein) %>%
    summarise(StdDev = sd(Precursor.Quantity), Avg = mean(Precursor.Quantity), CV = (StdDev/Avg)*100)

# CV filter, keep only those precursors that have CV < 20%
sample_matrix_small_cv_filt <- filter(sample_matrix_small_cv, CV > 20)

# sample_matrix_small_cv_filt_subset is the subset with precursors that have < CV 20%, this will be used for plots with CV < 20%
sample_matrix_small_cv_filt_subset <- filter(sample_matrix_small_20on20_final,!Protein %in% sample_matrix_small_cv_filt$Protein)

# save table with CV < 20%
#write.table(sample_matrix_small_cv_filt_subset, file = "sample_matrix_small_filtCV.tsv", sep="\t", row.names=F, quote=F) # export to file


```

## Precursor Counts {.tabset}
```{r count-precursors, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# All data
summarize_precursor <- sample_matrix_small %>% group_by(Category) %>% summarise(Count = n_distinct(Protein))

# Filtered 20on20
summarize_precursor_20on20_filt <- sample_matrix_small_20on20_final %>% group_by(Category) %>% summarise(Count = n_distinct(Protein))

# Filtered CV and 20on20
summarize_precursor_cv_filt <- sample_matrix_small_cv_filt_subset %>% group_by(Category) %>% summarise(Count = n_distinct(Protein))

```

### Before Filtering
```{r bar plot before filt, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

ggplot(data=summarize_precursor, aes(x=Count, y=Category, fill=Category)) +
  geom_bar(stat="identity") + 
  scale_fill_brewer(palette="Blues") + 
  labs(x = "Precursor Count",
      y = "Category") + 
  ggtitle("Precursor Count per Category") + 
  theme(plot.title = element_text(hjust = 0.5))

```

### After 20on20 Filtering
```{r bar plot after filt, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

ggplot(data=summarize_precursor_20on20_filt, aes(x=Count, y=Category, fill=Category)) +
  geom_bar(stat="identity") + 
  scale_fill_brewer(palette="Blues") + 
  labs(x = "Precursor Count",
      y = "Category") +
  ggtitle("Precursor Count per Category") + 
  theme(plot.title = element_text(hjust = 0.5))

```

### After CV and 20on20 Filtering
```{r bar plot after CV filt, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

ggplot(data=summarize_precursor_cv_filt, aes(x=Count, y=Category, fill=Category)) +
  geom_bar(stat="identity") + 
  scale_fill_brewer(palette="Blues") + 
  labs(x = "Precursor Count",
      y = "Category") +
  ggtitle("Precursor Count per Category") + 
  theme(plot.title = element_text(hjust = 0.5))

```


## Coefficient of Variation {.tabset}
```{r CV, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
#slide 11 - CV calculations  
# All data
cv_precursor <- sample_matrix_small %>% 
    group_by(Category,Protein) %>%
    summarise(StdDev = sd(Precursor.Quantity), Avg = mean(Precursor.Quantity), CV = StdDev/Avg*100)

# Filtered 20on20
cv_precursor_filt <- sample_matrix_small_20on20_final %>% 
    group_by(Category,Protein) %>%
    summarise(StdDev = sd(Precursor.Quantity), Avg = mean(Precursor.Quantity), CV = (StdDev/Avg)*100)

# Filtered CV and 20on20
cv_precursor_cv_filt <- sample_matrix_small_cv_filt_subset %>% 
    group_by(Category,Protein) %>%
    summarise(StdDev = sd(Precursor.Quantity), Avg = mean(Precursor.Quantity), CV = (StdDev/Avg)*100)
```

### Before Filtering
```{r violin before filt, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

p <- ggplot(cv_precursor, aes(x=Category, y=CV, fill=Category)) +
    geom_violin() + 
    scale_fill_brewer(palette="Blues") + 
    theme_classic() +
    stat_summary(fun.data="mean_sdl", mult=1, 
                 geom="pointrange", width=0.2 ) + 
    ggtitle("CV % per Category") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    scale_y_continuous(limits=c(0,200), breaks = scales::breaks_width(20)) 
p

```


### After 20on20 Filtering
```{r violin after filt, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

p <- ggplot(cv_precursor_filt, aes(x=Category, y=CV, fill=Category)) +
    geom_violin() + 
    scale_fill_brewer(palette="Blues") + 
    theme_classic() +
    stat_summary(fun.data="mean_sdl", mult=1, 
                 geom="pointrange", width=0.2 ) + 
    ggtitle("CV % per Category") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    scale_y_continuous(limits=c(0,200), breaks = scales::breaks_width(20)) 
p

```

### After CV and 20on20 Filtering
```{r violin after CV filt, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

p <- ggplot(cv_precursor_cv_filt, aes(x=Category, y=CV, fill=Category)) +
    geom_violin() + 
    scale_fill_brewer(palette="Blues") + 
    theme_classic() +
    stat_summary(fun.data="mean_sdl", mult=1, 
                 geom="pointrange", width=0.2 ) + 
    ggtitle("CV % per Category") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    scale_y_continuous(limits=c(0,200), breaks = scales::breaks_width(20)) 
p

```



## Precursor Abundance (Observed vs Expected) {.tabset}
```{r precursor abundance, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# slide 12 Precursor Abundance
# All data
median_grp_precursor = sample_matrix_small %>% group_by(Category, Protein) %>% summarise(Precursor.Quantity = median(Precursor.Quantity))

median_grp = median_grp_precursor %>% group_by(Category) %>%  summarise(Precursor.Quantity = median(Precursor.Quantity))

mean_grp = median_grp_precursor %>% group_by(Category) %>%  summarise(Precursor.Quantity = mean(Precursor.Quantity))

median_grp$median_percentage = (median_grp$Precursor.Quantity/median_grp$Precursor.Quantity[[1]]) * 100
median_grp$mean_percentage = (mean_grp$Precursor.Quantity/mean_grp$Precursor.Quantity[[1]]) * 100
#median_grp = head(median_grp, -1)
median_grp = median_grp[order(median_grp$Category, decreasing = TRUE),]
median_grp$expected = c(5, 10, 50, 100)
median_grp$scale = c(5, 10, 50, 100)

# Filtered 20on20
median_grp_precursor_filt = sample_matrix_small_20on20_final %>% group_by(Category, Protein) %>% summarise(Precursor.Quantity = median(Precursor.Quantity))

median_grp_filt = median_grp_precursor_filt %>% group_by(Category) %>%  summarise(Precursor.Quantity = median(Precursor.Quantity))

mean_grp_filt = median_grp_precursor_filt %>% group_by(Category) %>%  summarise(Precursor.Quantity = mean(Precursor.Quantity))

median_grp_filt$median_percentage = (median_grp_filt$Precursor.Quantity/median_grp_filt$Precursor.Quantity[[1]]) * 100
median_grp_filt$mean_percentage = (mean_grp_filt$Precursor.Quantity/mean_grp_filt$Precursor.Quantity[[1]]) * 100
median_grp_filt = median_grp_filt[order(median_grp_filt$Category, decreasing = TRUE),]
median_grp_filt$expected = c(5, 10, 50, 100)
median_grp_filt$scale = c(5, 10, 50, 100)

# Filtered CV < 20% and 20on20
median_grp_precursor_cv_filt = sample_matrix_small_cv_filt_subset %>% group_by(Category, Protein) %>% summarise(Precursor.Quantity = median(Precursor.Quantity))

median_grp_cv_filt = median_grp_precursor_cv_filt %>% group_by(Category) %>%  summarise(Precursor.Quantity = median(Precursor.Quantity))

mean_grp_cv_filt = median_grp_precursor_cv_filt %>% group_by(Category) %>%  summarise(Precursor.Quantity = mean(Precursor.Quantity))

median_grp_cv_filt$median_percentage = (median_grp_cv_filt$Precursor.Quantity/median_grp_cv_filt$Precursor.Quantity[[1]]) * 100
median_grp_cv_filt$mean_percentage = (mean_grp_cv_filt$Precursor.Quantity/mean_grp_cv_filt$Precursor.Quantity[[1]]) * 100
median_grp_cv_filt = median_grp_cv_filt[order(median_grp_cv_filt$Category, decreasing = TRUE),]
median_grp_cv_filt$expected = c(5, 10, 50, 100)
median_grp_cv_filt$scale = c(5, 10, 50, 100)
```

### Before Filtering
```{r line graph before, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# All data
colors <- c("Expected" = "black", "Median %" = "red", "Mean %" = "steelblue")

ggplot(median_grp, aes(scale)) +
  scale_color_manual(values = colors) +
  geom_line(aes(y = expected, color = "Expected")) +
  geom_point(aes(y = expected, color = "Expected")) +
  geom_line(aes(y = median_percentage, color = "Median %")) +
  geom_point(aes(y = median_percentage, color = "Median %")) +
  geom_line(aes(y = mean_percentage, color = "Mean %")) +
  geom_point(aes(y = mean_percentage, color = "Mean %")) +
  scale_x_discrete(limits=c(5,10,50,100)) +
  expand_limits(x=0, y=0) +
  #scale_x_continuous(limits=c(0,100)) +
  scale_y_continuous(limits=c(0,100), breaks = scales::breaks_width(10)) +
  labs(x = "Concentration",
        y = "Precursor abundance (%)",
        color = "Legend") + 
  ggtitle("Precursor abundance across concentrations") +
    theme(plot.title = element_text(hjust = 0.5))
```

### After 20on20 Filtering
```{r line graph after, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# Filtered
colors <- c("Expected" = "black", "Median %" = "red", "Mean %" = "steelblue")

ggplot(median_grp_filt, aes(scale)) +
  scale_color_manual(values = colors) +
  geom_line(aes(y = expected, color = "Expected")) +
  geom_point(aes(y = expected, color = "Expected")) +
  geom_line(aes(y = median_percentage, color = "Median %")) +
  geom_point(aes(y = median_percentage, color = "Median %")) +
  geom_line(aes(y = mean_percentage, color = "Mean %")) +
  geom_point(aes(y = mean_percentage, color = "Mean %")) +
  scale_x_discrete(limits=c(5,10,50,100)) +
  expand_limits(x=0, y=0) +
  #scale_x_continuous(limits=c(0,100)) +
  scale_y_continuous(limits=c(0,100), breaks = scales::breaks_width(10)) +
  labs(x = "Concentration",
        y = "Precursor abundance (%)",
        color = "Legend") + 
  ggtitle("Precursor abundance across concentrations, 20on20") +
    theme(plot.title = element_text(hjust = 0.5))

```

### After CV and 20on20 Filtering
```{r line graph after CV, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# Filtered
colors <- c("Expected" = "black", "Median %" = "red", "Mean %" = "steelblue")

ggplot(median_grp_cv_filt, aes(scale)) +
  scale_color_manual(values = colors) +
  geom_line(aes(y = expected, color = "Expected")) +
  geom_point(aes(y = expected, color = "Expected")) +
  geom_line(aes(y = median_percentage, color = "Median %")) +
  geom_point(aes(y = median_percentage, color = "Median %")) +
  geom_line(aes(y = mean_percentage, color = "Mean %")) +
  geom_point(aes(y = mean_percentage, color = "Mean %")) +
  scale_x_discrete(limits=c(5,10,50,100)) +
  expand_limits(x=0, y=0) +
  #scale_x_continuous(limits=c(0,100)) +
  scale_y_continuous(limits=c(0,100), breaks = scales::breaks_width(10)) +
  labs(x = "Concentration",
        y = "Precursor abundance (%)",
        color = "Legend") + 
  ggtitle("Precursor abundance across concentrations, CV<20% & 20on20") +
    theme(plot.title = element_text(hjust = 0.5))

```


## Difference in Distribution of Retention Time {.tabset}
```{r RT, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
# All data
rt_precursor <- sample_matrix_small %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(RT))

# Filtered 20on20
rt_precursor_filt <- sample_matrix_small_20on20_final %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(RT))

# Filtered CV and 20on20
rt_precursor_cv_filt <- sample_matrix_small_cv_filt_subset %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(RT))
```

### Before Filtering
```{r before filtering RT, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
for (i in c("B", "C", "D")) {
  sub_A <- filter(rt_precursor, Category == "A")
  sub_i <- filter(rt_precursor, Category == i)
  
  merged_A_i <- merge(x = sub_A, y = sub_i, by = "Protein")
  merged_A_i <- merged_A_i %>% mutate(diff_iA = Median.y-Median.x)
  
  p <- ggplot(merged_A_i, aes(diff_iA)) +
  geom_histogram(binwidth = 0.01, color="steelblue", fill="steelblue") +
  ggtitle(paste0("RT difference between ", i, " and A")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") + 
  ylab("Frequency") + 
  xlim(-1.5, 1.5)
  
  print(p)
}
```

### After 20on20 Filtering
```{r after filetring RT, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
for (i in c("B", "C", "D")) {
  sub_A <- filter(rt_precursor_filt, Category == "A")
  sub_i <- filter(rt_precursor_filt, Category == i)
  
  merged_A_i <- merge(x = sub_A, y = sub_i, by = "Protein")
  merged_A_i <- merged_A_i %>% mutate(diff_iA = Median.y-Median.x)
  
  p <- ggplot(merged_A_i, aes(diff_iA)) +
  geom_histogram(binwidth = 0.01, color="steelblue", fill="steelblue") +
  ggtitle(paste0("RT difference between ", i, " and A")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") + 
  ylab("Frequency") + 
  xlim(-1.5, 1.5)
  
  print(p)
}

```

### After CV and 20on20 Filtering
```{r after CV filetring RT, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
for (i in c("B", "C", "D")) {
  sub_A <- filter(rt_precursor_cv_filt, Category == "A")
  sub_i <- filter(rt_precursor_cv_filt, Category == i)
  
  merged_A_i <- merge(x = sub_A, y = sub_i, by = "Protein")
  merged_A_i <- merged_A_i %>% mutate(diff_iA = Median.y-Median.x)
  
  p <- ggplot(merged_A_i, aes(diff_iA)) +
  geom_histogram(binwidth = 0.01, color="steelblue", fill="steelblue") +
  ggtitle(paste0("RT difference between ", i, " and A")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") + 
  ylab("Frequency") + 
  xlim(-1.5, 1.5)
  
  print(p)
}

```

## Difference in Distribution of Ion Mobility {.tabset}
```{r IM, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
#All data
im_precursor <- sample_matrix_small %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(Exp.1.K0))

# Filtered 20on20
im_precursor_filt <- sample_matrix_small_20on20_final %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(Exp.1.K0))

# Filtered CV and 20on20
im_precursor_cv_filt <- sample_matrix_small_cv_filt_subset %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(Exp.1.K0))
```

### Before Filtering
```{r before filtering IM, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
for (i in c("B", "C", "D")) {
  sub_A <- filter(im_precursor, Category == "A")
  sub_i <- filter(im_precursor, Category == i)
  
  merged_A_i <- merge(x = sub_A, y = sub_i, by = "Protein")
  merged_A_i <- merged_A_i %>% mutate(diff_iA = Median.y-Median.x)
  
  p <- ggplot(merged_A_i, aes(diff_iA)) +
  geom_histogram(binwidth = 0.0001, color="steelblue", fill="steelblue") +
  ggtitle(paste0("IM difference between ", i, " and A")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") + 
  ylab("Frequency") + 
  xlim(-0.5, 0.5)
  
  print(p)
}
```

### After 20on20 Filtering
```{r after filtering IM, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
for (i in c("B", "C", "D")) {
  sub_A <- filter(im_precursor_filt, Category == "A")
  sub_i <- filter(im_precursor_filt, Category == i)
  
  merged_A_i <- merge(x = sub_A, y = sub_i, by = "Protein")
  merged_A_i <- merged_A_i %>% mutate(diff_iA = Median.y-Median.x)
  
  p <- ggplot(merged_A_i, aes(diff_iA)) +
  geom_histogram(binwidth = 0.0001, color="steelblue", fill="steelblue") +
  ggtitle(paste0("IM difference between ", i, " and A")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") + 
  ylab("Frequency") + 
  xlim(-0.5, 0.5)
  
  print(p)
}

```

### After CV and 20on20 Filtering
```{r after CV filtering IM, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
for (i in c("B", "C", "D")) {
  sub_A <- filter(im_precursor_cv_filt, Category == "A")
  sub_i <- filter(im_precursor_cv_filt, Category == i)
  
  merged_A_i <- merge(x = sub_A, y = sub_i, by = "Protein")
  merged_A_i <- merged_A_i %>% mutate(diff_iA = Median.y-Median.x)
  
  p <- ggplot(merged_A_i, aes(diff_iA)) +
  geom_histogram(binwidth = 0.0001, color="steelblue", fill="steelblue") +
  ggtitle(paste0("IM difference between ", i, " and A")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("") + 
  ylab("Frequency") + 
  xlim(-0.5, 0.5)
  
  print(p)
}

```

## Distribution of Precursors based on Correlation {.tabset}
```{r prec with good correlation, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}


# Subset to get good, bad and ugly precursors based on correlation
paser_small_corr_good <- filter(paser_small_corr_file, Corr >= 0.9)
paser_small_corr_bad <- filter(paser_small_corr_file, between(Corr, 0.5, 0.7))
paser_small_corr_ugly <- filter(paser_small_corr_file, between(Corr, 0.0, 0.4))

```

### Good
```{r Good, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

ggplot(paser_small_corr_good, aes(x=Category, y=log2(Precursor.Quantity), fill=Category)) +
  geom_boxplot() +
  scale_fill_brewer(palette="Blues") +
  theme(
    legend.position="none",
    plot.title = element_text(hjust = 0.5)
  ) +
  ggtitle("Distribution of Precursors across Concentrations - Good") +
  labs(x = "Concentrations",
      y = "Precursor abundance (log2)")

```

### Bad
```{r prec with bad correlation, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

ggplot(paser_small_corr_bad, aes(x=Category, y=log2(Precursor.Quantity), fill=Category)) +
  geom_boxplot() +
  scale_fill_brewer(palette="Blues") +
  theme(
    legend.position="none",
    plot.title = element_text(hjust = 0.5)
  ) +
  ggtitle("Distribution of Precursors across Concentrations - Bad") +
  labs(x = "Concentrations",
      y = "Precursor abundance (log2)")

```

### Ugly
```{r prec with ugly correlation, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

ggplot(paser_small_corr_ugly, aes(x=Category, y=log2(Precursor.Quantity), fill=Category)) +
  geom_boxplot() +
  scale_fill_brewer(palette="Blues") +
  theme(
    legend.position="none",
    plot.title = element_text(hjust = 0.5)
  ) +
  ggtitle("Distribution of Precursors across Concentrations - Ugly") +
  labs(x = "Concentrations",
      y = "Precursor abundance (log2)")

```

```{r getting prec subsets for good, bad and ugly, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
#manually assigned list of good precursors based on investigation

good_prec = c('Q08043-EGLLLWC(UniMod:4)QR2','P62701-YALTGDEVKK2','P13639-ETVSEESNVLC(UniMod:4)LSK2','P83731-QINWTVLYR2','P60174-IAVAAQNC(UniMod:4)YK2')

#filtering based on precursors
prec_good_list = filter(paser_small_corr_file, Protein %in% good_prec)

prec_good_list <- prec_good_list %>%
  group_by(Category,Protein)  %>%
  summarise(median = median(Precursor.Quantity))  %>%
  group_by(Protein) %>%
  mutate(Col = median/median[Category == 'A']*100)

#ordering based on category
prec_good_list = prec_good_list[order(prec_good_list$Category, decreasing = TRUE),]

#manually assigned list of bad precursors based on investigation. 
bad_prec <- c("Q00610-IHEGC(UniMod:4)EEPATHNALAK3","P06576-IMDPNIVGSEHYDVAR2","P14854-NC(UniMod:4)WQNYLDFHR3","P11387-HLQDLMEGLTAK2","P12236-DFLAGGIAAAISK2")

#filtering based on precursors
prec_bad_list = filter(paser_small_corr_file, Protein %in% bad_prec)

prec_bad_list <- prec_bad_list %>%
  group_by(Category,Protein)  %>%
  summarise(median = median(Precursor.Quantity))  %>%
  group_by(Protein) %>%
  mutate(Col = median/median[Category == 'A']*100)

#ordering based on category
prec_bad_list = prec_bad_list[order(prec_bad_list$Category, decreasing = TRUE),]

#manually assigned list of bad precursors based on investigation. 
ugly_prec <- c("P49419-GEVITTYC(UniMod:4)PANNEPIAR2","Q96PK6-IFVGNVSAAC(UniMod:4)TSQELR2","P04843-IDHILDAL2","Q96QK1-SEDPDQQYLILNTAR2","P30050-HPHDIIDDINSGAVEC(UniMod:4)PAS2")

#filtering based on precursors
prec_ugly_list = filter(paser_small_corr_file, Protein %in% ugly_prec)

prec_ugly_list <- prec_ugly_list %>%
  group_by(Category,Protein)  %>%
  summarise(median = median(Precursor.Quantity))  %>%
  group_by(Protein) %>%
  mutate(Col = median/median[Category == 'A']*100)

#ordering based on precursors
prec_ugly_list = prec_ugly_list[order(prec_ugly_list$Category, decreasing = TRUE),]

```

## Observed vs Expected for Specific Precursors {.tabset}
### Good
```{r line good, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

#adding the concentration scale
prec_good_list <- prec_good_list %>%
  mutate(scale = case_when(
    grepl(pattern = "A", x = Category) ~ 100,
    grepl(pattern = "B", x = Category) ~ 50,
    grepl(pattern = "C", x = Category) ~ 10,
    grepl(pattern = "D", x = Category) ~ 5,
  ))

ggplot(prec_good_list, aes(scale, group=Protein)) +
  #scale_color_manual(values = colors) +
  geom_line(aes(y = Col, color=Protein)) +
  geom_point(aes(y = Col, color=Protein)) +
  expand_limits(x=0, y=0) +
  scale_y_continuous(limits=c(0,120), breaks = scales::breaks_width(10)) +
  labs(x = "Concentration",
       y = "Precursor abundance (%)",
       color = "Protein") +
  ggtitle("Precursor abundance across concentrations") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Bad
```{r line bad, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

#adding the concentration scale
prec_bad_list <- prec_bad_list %>%
  mutate(scale = case_when(
    grepl(pattern = "A", x = Category) ~ 100,
    grepl(pattern = "B", x = Category) ~ 50,
    grepl(pattern = "C", x = Category) ~ 10,
    grepl(pattern = "D", x = Category) ~ 5,
  ))

ggplot(prec_bad_list, aes(scale, group=Protein)) +
  #scale_color_manual(values = colors) +
  geom_line(aes(y = Col, color=Protein)) +
  geom_point(aes(y = Col, color=Protein)) +
  expand_limits(x=0, y=0) +
  scale_y_continuous(limits=c(0,120), breaks = scales::breaks_width(10)) +
  labs(x = "Concentration",
       y = "Precursor abundance (%)",
       color = "Protein") +
  ggtitle("Precursor abundance across concentrations") +
  theme(plot.title = element_text(hjust = 0.5))

```

### Ugly
```{r line ugly, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

#adding the concentration scale
prec_ugly_list <- prec_ugly_list %>%
  mutate(scale = case_when(
    grepl(pattern = "A", x = Category) ~ 100,
    grepl(pattern = "B", x = Category) ~ 50,
    grepl(pattern = "C", x = Category) ~ 10,
    grepl(pattern = "D", x = Category) ~ 5,
  ))

ggplot(prec_ugly_list, aes(scale, group=Protein)) +
  #scale_color_manual(values = colors) +
  geom_line(aes(y = Col, color=Protein)) +
  geom_point(aes(y = Col, color=Protein)) +
  expand_limits(x=0, y=0) +
  scale_y_continuous(limits=c(0,120), breaks = scales::breaks_width(10)) +
  labs(x = "Concentration",
       y = "Precursor abundance (%)",
       color = "Protein") +
  ggtitle("Precursor abundance across concentrations") +
  theme(plot.title = element_text(hjust = 0.5))

```

## RT for Specific Precursors {.tabset}
### Good
```{r prec with good RT, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

#subsetting based on the paser correlation file and good precursors
good_prec_df <- subset(paser_small_corr_file, Protein %in% good_prec)

good_prec_median <- good_prec_df %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(RT))

ggplot(good_prec_median, aes(x= Category, y = Median, group=Protein)) +
  geom_line(aes(color=Protein)) +
  geom_point(aes(color=Protein)) + 
  ylim(0, 25) + 
  labs(x="Category", y="RT (min)") +
  ggtitle("RT for Precursors with Good Regression") +
  theme(plot.title = element_text(hjust = 0.5))

```

### Bad
```{r prec with bad RT, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

#subsetting based on the paser correlation file and bad precursors
bad_prec_df <- subset(paser_small_corr_file, Protein %in% bad_prec)

bad_prec_median <- bad_prec_df %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(RT))

ggplot(bad_prec_median, aes(x= Category, y = Median, group=Protein)) +
  geom_line(aes(color=Protein)) +
  geom_point(aes(color=Protein)) + 
  ylim(0, 25) +
  labs(x="Category", y="RT (min)") +
  ggtitle("RT for Precursors with Bad Regression") +
  theme(plot.title = element_text(hjust = 0.5))

```

### Ugly
```{r prec with ugly RT, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

#subsetting based on the paser correlation file and ugly precursors
ugly_prec_df <- subset(paser_small_corr_file, Protein %in% ugly_prec)

ugly_prec_median <- ugly_prec_df %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(RT))

ggplot(ugly_prec_median, aes(x= Category, y = Median, group=Protein)) +
  geom_line(aes(color=Protein)) +
  geom_point(aes(color=Protein)) + 
  ylim(0, 25) + 
  labs(x="Category", y="RT (min)") +
  ggtitle("RT for Precursors with Ugly Regression") +
  theme(plot.title = element_text(hjust = 0.5))

```

## IM for Specific Precursors {.tabset}
### Good
```{r prec with good IM, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}

good_prec_median_IM <- good_prec_df %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(Exp.1.K0))

ggplot(good_prec_median_IM, aes(x= Category, y = Median, group=Protein)) +
  geom_line(aes(color=Protein)) +
  geom_point(aes(color=Protein)) + 
  ylim(0, 2) + 
  labs(x="Category", y="Mobility (1/k0)") +
  ggtitle("IM for Precursors with Good Regression") +
  theme(plot.title = element_text(hjust = 0.5))

```

### Bad
```{r prec with bad IM, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
bad_prec_median_IM <- bad_prec_df %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(Exp.1.K0))

ggplot(bad_prec_median_IM, aes(x= Category, y = Median, group=Protein)) +
  geom_line(aes(color=Protein)) +
  geom_point(aes(color=Protein)) + 
  ylim(0, 2) +
  labs(x="Category", y="Mobility (1/k0)") +
  ggtitle("IM for Precursors with Good Regression") +
  theme(plot.title = element_text(hjust = 0.5))

```

### Ugly
```{r prec with ugly IM, echo=F, results='asis', fig.keep='all', message=FALSE, warning=FALSE}
ugly_prec_median_IM <- ugly_prec_df %>% 
    group_by(Category,Protein) %>%
    summarise(Median = median(Exp.1.K0))

ggplot(ugly_prec_median_IM, aes(x= Category, y = Median, group=Protein)) +
  geom_line(aes(color=Protein)) +
  geom_point(aes(color=Protein)) + 
  ylim(0, 2) + 
  labs(x="Category", y="Mobility (1/k0)") +
  ggtitle("IM for Precursors with Good Regression") +
  theme(plot.title = element_text(hjust = 0.5))

```
